Trajectories

### 2022-08-29 docs/index.Rmd
Update report

---
title: "Trajectories"
output:
  github_document:
    pandoc_args: --webtex
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    message = TRUE,
    echo = FALSE,
    warning = TRUE,
    include = TRUE,
    out.width = "80%",
    dpi = 250
)
options(
    knitr.kable.NA = "-",
    knitr.duplicate.label = "allow",
    ggplot2.discrete.fill = ggsci::pal_d3()(4),
    ggplot2.discrete.colour = ggsci::pal_d3()(4),
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# set params
tar_load_globals()
tar_load_all()
theme_set(theme_custom())

@@

@@

```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies. Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}

items %>%
    left_join(select(multilex_data$pool, te, language)) %>% 
    unnest(list) %>% 
    group_by(language, list) %>% 
    summarise_at(
        vars(freq, n_phon, lv), 
        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
    ) %>% 
    select(-c(n_phon_n, lv_n)) %>% 
    rename(n = freq_n) %>% 
    gt(groupname_col = "language") %>% 
    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
    fmt_percent(matches("lv_")) %>% 
    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
    cols_label(
        list = "",
        n = md("*N*"),
        freq_mean = md("Mean"),
        freq_sd = md("*SD*"),
        freq_min = md("Range"),
        n_phon_mean = md("Mean"),
        n_phon_sd = md("*SD*"),
        n_phon_min = md("Range"),
        lv_mean = md("Mean"),
        lv_sd = md("*SD*"),
        lv_min = md("Range")
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(everything())
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(everything())
    )

@@

```{r items-frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>% 
    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>% 
    ggplot() +
    aes(
        x = freq,
        fill = language,
        colour = language,
        shape = language
    ) +
    stat_slab(
        slab_colour = "white",
        slab_size = 1,
        alpha = 0.5,
    ) +
    stat_pointinterval(
        aes(y = -0.1),
        position = position_dodge(width = 0.1),
        point_size = 3
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "Frequency (Zipf)",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
    theme(
        legend.position = "top",
        axis.title = element_blank()
    )

@@

```{r items-nphon, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>%
    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>% 
    count(language, n_phon) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(
        x = n_phon,
        y = prop,
        fill = language,
        colour = language,
        shape = language
    ) +
    geom_col(
        position = position_dodge()
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "Frequency (Zipf)",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
    theme(
        legend.position = "top",
        axis.title = element_blank()
    )

@@

```{r items-lv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv_cut) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(x = lv_cut, y = prop) +
    geom_col(fill = ggsci::pal_d3()(4)[1]) +
    labs(
        x = "Levenshtein",
        y = "Probability density",
    ) +
    theme(
        legend.position = "top",
        axis.title = element_blank()
    )

@@

```{r items-frequency-phonemes, message=FALSE, warning=FALSE, paged.print=FALSE}

items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    )

@@

```{r items-frequency-levenshtein, message=FALSE, warning=FALSE, paged.print=FALSE}

items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(freq, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.5) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )

@@

```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
    mutate(dominant_language = ifelse(doe_catalan >= doe_spanish, "Catalan", "Spanish")) %>% 
    group_by(time_stamp, dominant_language) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    group_by(dominant_language) %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
    guides(colour = guide_legend(ncol = 2)) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
    )

@@

```{r participants-age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col() +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()

@@

participants %>% 
    mutate(
        dominant_language = ifelse(doe_catalan >= doe_spanish, "Catalan", "Spanish"),
        age = round(age),
        id = paste(id, "(", time, ")"),
        rank = rank(doe_catalan)
    ) %>% 
    pivot_longer(
        starts_with("doe_"), 
        names_to = "language",
        values_to = "doe",
        names_prefix = "doe_",
        names_transform = list(language = str_to_sentence)
    ) %>% 
    ggplot() +
    aes(x = doe, y = reorder(id, -rank), fill = language) +
    geom_col(width = 0.75, position = position_fill(), colour = "white", size = 0.1) +
    labs(
        x = "Degree of exposure to Catalan",
        y = "Participant",
        colour = "Language",
        fill = "Language"
    ) +
    scale_x_continuous(labels = percent) +
    theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()
    )

@@

```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>%
    left_join(select(multilex_data$logs, id, time, edu_parent)) %>% 
    mutate(
        edu_parent = fct_explicit_na(as.factor(edu_parent)),
        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
    ) %>% 
    count(doe_catalan, edu_parent) %>% 
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(multilex_data$logs$edu_parent)
        )
    ) %>% 
    rename(total = n) %>% 
    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]))) %>% 
    rename(Total = total) %>% 
    select(-"NA") %>% 
    gt() %>% 
    cols_label(
        doe_catalan = "DoE Catalan"
    ) %>% 
    summary_rows(
        columns = 2:7, 
        fns = list(Total = ~sum(., na.rm = TRUE)),
        formatter = fmt_number,
        decimals = 0
    ) %>% 
    fmt_missing(columns = 3:8, missing_text = "--") %>% 
    tab_spanner("Educational attainment", 3:8) %>% 
    tab_style(
        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
        locations = cells_body(columns = 9, rows = 1:10)
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

@@

We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_std` by `item`, and their correlation parameter. We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope. Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`. The models were implemented in `brms` as:

* `understands/produces ~ 1 + age_std + frequency + (1 + age_std | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age_std + frequency + doe + (1 + age_std + doe | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age_std + frequency + doe*cognate + (1 + age_std + doe | item) + (1 + frequency + cognate| id)`

@@

$$
\begin{aligned}
\color{white}\log(\frac{p}{n - q}) &= \color{white}(\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
& \color{white}(\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
& \color{white}(\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
& \color{white}(\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
& \color{white}(\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
& \color{white}(\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
& \color{white}\varepsilon_{pi}
\end{aligned}
$$

@@

```{stan, eval = FALSE, tidy = FALSE}

@@

```{r loos, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, eval = FALSE}
loo_comp %>% 
    as_tibble() %>% 
    rownames_to_column("model") %>% 
    relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
    mutate_all(as.numeric) %>% 
    gt() %>% 
    tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
    tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
    tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
    tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
    fmt_number(3:13) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("SE"),
        subsampling_se_p_loo = md("SE<sub>sub</sub>"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        subsampling_se_looic = md("SE<sub>sub</sub>"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*"),
        subsampling_se_diff = md("*SE<sub>sub</sub>*"),
    ) %>% 
    as_raw_html()
```

@@

```{r fix_effects-plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

>>>>

---
title: "Trajectories"
authors:
    - Gonzalo Garcia-Castro^[Universitat Pompeu Fabra, gonzalo.garciadecastro@upf.edu]
    - Daniela Avila-Varela^[Universitat Pompeu Fabra]
    - Ignacio Castillejo^[Universidad Autónoma de Madrid]
    - Núria Sebastian-Galles^[Universitat Pompeu Fabra]
output:
    html_document:
        code_folding: hide
        number_sections: true
        toc: true
    github_document:
        pandoc_args: --webtex
        number_sections: true
        toc: true
bibliography: "../references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    message = TRUE,
    echo = TRUE,
    warning = TRUE,
    include = TRUE,
    out.width = "80%",
    dpi = 250
)
options(
    knitr.kable.NA = "-",
    knitr.duplicate.label = "allow",
    ggplot2.discrete.fill = ggsci::pal_d3()(4),
    ggplot2.discrete.colour = ggsci::pal_d3()(4),
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

```{r prepare, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# set params and load target contents
tar_load_globals()
tar_load_all()
theme_set(theme_custom())

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

@@

```{r items, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
pool <- multilex::pool %>% 
    drop_na(cognate)
n_categories <- distinct(pool, category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
n_item_cognate <- count(pool, language, cognate) %>% 
    pull(n) %>% 
    set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
```

@@

```{r items_frequency, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies. Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}

items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))

items %>%
    unnest(list) %>% 
    group_by(language, list) %>% 
    summarise_at(
        vars(freq, n_phon, lv), 
        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
    ) %>% 
    select(-c(n_phon_n, lv_n)) %>% 
    rename(n = freq_n) %>% 
    gt(groupname_col = "language") %>% 
    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
    fmt_percent(matches("lv_")) %>% 
    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
    cols_label(
        list = "",
        n = md("*N*"),
        freq_mean = md("Mean"),
        freq_sd = md("*SD*"),
        freq_min = md("Range"),
        n_phon_mean = md("Mean"),
        n_phon_sd = md("*SD*"),
        n_phon_min = md("Range"),
        lv_mean = md("Mean"),
        lv_sd = md("*SD*"),
        lv_min = md("Range")
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(everything())
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(everything())
    )

@@

```{r items-frequency, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>% 
    ggplot() +
    aes(
        x = freq,
        fill = language,
        colour = language,
        shape = language
    ) +
    stat_slab(
        slab_colour = "white",
        slab_size = 1,
        alpha = 0.5,
    ) +
    stat_pointinterval(
        aes(y = -0.1),
        position = position_dodge(width = 0.1),
        point_size = 3
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "Frequency (Zipf)",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
    theme(
        legend.position = "top",
        legend.title = element_blank()
    )

@@

```{r items-nphon, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>%
    count(language, n_phon) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(
        x = n_phon,
        y = prop,
        fill = language,
        colour = language,
        shape = language
    ) +
    geom_col(
        position = position_dodge(width = 0.95)
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "# Phonemes",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    scale_x_continuous(breaks = seq(0, 20, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(),
        legend.title = element_blank()
    )

@@

```{r items-lv, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv_cut) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(x = lv_cut, y = prop) +
    geom_col(fill = ggsci::pal_d3()(4)[1]) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    labs(
        x = "Levenshtein",
        y = "Probability density",
    ) +
    theme(
        legend.position = "top"
    )

@@

```{r items-frequency-phonemes, message=FALSE, warning=FALSE, paged.print=FALSE}
items %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r items-frequency-levenshtein, message=FALSE, warning=FALSE, paged.print=FALSE}
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(freq, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.5) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
participants <- left_join(
    participants, 
    select(multilex_data$logs, id, time, edu_parent, dominant_language = dominance)
)

participants %>% 
    group_by(time_stamp, dominant_language) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    group_by(dominant_language) %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
    guides(colour = guide_legend(ncol = 2)) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
    )

@@

```{r participants-age, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = pal_d3()(3)[3]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()

@@

participants %>% 
    mutate(
        age = round(age),
        id = paste(id, "(", time, ")"),
        rank = rank(doe_catalan)
    ) %>% 
    pivot_longer(
        starts_with("doe_"), 
        names_to = "language",
        values_to = "doe",
        names_prefix = "doe_",
        names_transform = list(language = str_to_sentence)
    ) %>% 
    ggplot() +
    aes(x = doe, y = reorder(id, -rank), fill = language) +
    geom_col(width = 1, position = position_fill(), colour = "white", size = 0.1) +
    geom_vline(xintercept = seq(0, 1, 0.25), colour = "grey") +
    labs(
        x = "Degree of Exposure (DoE)",
        y = "Participant",
        colour = "Language",
        fill = "Language"
    ) +
    scale_x_continuous(labels = percent) +
    theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.title = element_blank()
    )

@@

```{r participants_edu, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>%
    mutate(
        edu_parent = fct_explicit_na(as.factor(edu_parent)),
        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
    ) %>% 
    count(doe_catalan, edu_parent) %>% 
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(multilex_data$logs$edu_parent)
        )
    ) %>% 
    rename(total = n) %>% 
    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
    rename(Total = total) %>% 
    select(-"NA") %>% 
    gt() %>% 
    cols_label(
        doe_catalan = "DoE Catalan"
    ) %>% 
    summary_rows(
        columns = 2:7, 
        fns = list(Total = ~sum(., na.rm = TRUE)),
        formatter = fmt_number,
        decimals = 0
    ) %>% 
    fmt_missing(columns = 3:8, missing_text = "--") %>% 
    tab_spanner("Educational attainment", 3:8) %>% 
    tab_style(
        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
        locations = cells_body(columns = 9, rows = 1:10)
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

@@

We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_std` by `item`, and their correlation parameter. We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope. Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`.

@@

$$\begin{aligned}
g[\mathrm{Pr}(y_{ij} \le a_s)] = \tau_s - \nu_{ij}
\end{aligned}$$

$s$ is the s-th response category, $a_1 < a_2 < a_3$ are ordered response categories, $\mathrm{Pr}(y_i \le a_3) = 1$ and $\tau_s$ are threshold parameters where $\tau_1 < \tau_2 < \tau_3$.

$$
\begin{aligned}

\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} +
\beta_6\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_7\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_8\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + 
\beta_9{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) +\\

&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{TE})} \mathrm{Freq}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{5j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) +\\

&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{ID})} \mathrm{Freq}_{j} +   \theta_{3j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{5j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{6j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) + 
\theta_{7j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}) +
\theta_{8j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j})

\end{aligned}
$$

## Model prior

$$\begin{aligned}
\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
\beta_{1} &\sim \mathrm{Normal}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
\beta_{2-9} &\sim \mathrm{Normal}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]
\end{aligned}
$$

In addition, we specified a $LKJ$ prior ($\sim LKJ(2)$ for the correlations between group-level coefficients. We implemented the model prior following {brms} syntax as:

```r
c(
prior(normal(-0.25, 0.1), class = "Intercept"),
prior(normal(1, 0.1), class = "sd", group = "te"),
prior(normal(1, 0.1), class = "sd", group = "id"),
prior(lkj(2), class = "cor"),
prior(normal(1, 0.1), class = "b", coef = "age_std"),
prior(normal(0, 0.1), class = "b", coef = "freq_std"),
prior(normal(0, 0.1), class = "b", coef = "n_phon_std"),
prior(normal(0, 0.1), class = "b", coef = "doe_std"),
prior(normal(0, 0.1), class = "b", coef = "lv_std"),
prior(normal(0, 0.1), class = "b", coef = "doe_std:lv_std"),
prior(normal(0, 0.1), class = "b", coef = "age_std:doe_std"),
prior(normal(0, 0.1), class = "b", coef = "age_std:lv_std"),
prior(normal(0, 0.1), class = "b", coef = "age_std:doe_std:lv_std")
)
```

## Model settings

* OS: `r sessionInfo()$running`, `r `sessionInfo()$platform`
* R version: `r sessionInfo()$R.version$version.string`
* Algorithm: Hamiltonian Montecarlo (No U-turn Sampler, *NUTS*)
* Engine: Stan (brms interface, `r sessionInfo()$otherPkgs$brms$Version`)
* brms backend: CmdStanR (cmdstanr `r sessionInfo()$loadedOnly$cmdstanr$Version`)
* Chains: `r dim(model_fit_7$fit)[2]`
* Cores: `r dim(model_fit_7$fit)[2]`
* Iterations: `r dim(model_fit_7$fit)[1]` (`r dim(model_fit_7$fit)[1]/2`)

More details in Appendix: Session Info.

@@

@@

```{r loos, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
model_loos %>% 
    as_tibble() %>% 
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic"), 
        matches("diff")
    ) %>%
    mutate_all(as.numeric) %>% 
    gt() %>% 
    tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
    tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
    tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
    tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
    fmt_number(3:13) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("SE"),
        subsampling_se_p_loo = md("SE<sub>sub</sub>"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        subsampling_se_looic = md("SE<sub>sub</sub>"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*"),
        subsampling_se_diff = md("*SE<sub>sub</sub>*"),
    ) 
```

@@

```{r fix_effects-plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
var_repl <- c(
    "b_Intercept[1]" = "Intercept (Comprehension)",
    "b_Intercept[2]" = "Intercept (Production)",
    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein", 
    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
    "b_age_std:doe_std" = "Age \u00d7 DoE",
    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein", 
    "b_Intercept" = "Intercept", 
    "b_age_std" = "Age (+1 month)", 
    "b_freq_std" = "Frequency (+1 SD)",
    "b_n_phon_std" = "# Phonemes (+1 SD)",
    "b_doe_std" = "DoE (+1 SD)", 
    "b_lv_std" = "Levenshtein (+1 SD)"
)
